Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-04 Paper/paper.tex
revised 2.2

Since sliding window embeddings of periodic videos lie on a topological loop, this means that, at an appropriate scale, a graph built on the sliding window point cloud will be approximately circular.  We can then compute the sliding window phase $\phi \in [0, 2\pi]$ from this graph using the {\em graph Laplacian} \cite{chung1997spectral}, which we now describe. We first construct an adjacency matrix $\mb{A}$ in which the entry at $\mb{A}_{ij}$ represents the similarity between windows at times $i$ and $j$.  Given a scale $\sigma$, We define an unweighted version as $\mb{A}_{ij}$ is 1 if $||SW_{D}[\mb{Z}(i)] - SW_{D}[\mb{Z}(j)]||_2 \leq \sigma$ and 0 otherwise. We also define a weighted version in which $\mb{A}_{ij} = exp(-||SW_{D}[\mb{Z}(i)] - SW_{D}[\mb{Z}(j)]||_2/2\sigma^2)$. In both cases, we define the graph Laplacian as:
\begin{equation}
	\mb{L} = \mb{D}-\mb{A}
\end{equation}
where $\mb{D}_{ii} = \sum_{j = 1}^N \mb{A}_{ij}, \mb{D}_{i \neq j} = 0$ is the degree matrix representing the sum of all outgoing weights on the diagonal.

We now examine the unweighted Laplacain of an ideal model to motivate our approach.  Suppose we have performed a sliding window $SW_{D}[X[n]]$ for a video $X$ which repeats itself exactly every $T$ frames $k$ times total, for a total of $N = kT$ windows.  Let us further suppose it is possible to choose an $\sigma$ so that the unweighted adjacency matrix contains a 1 entry for the edge between window directly before and after it in time, as well as the windows directly before and after its repetitions, but none others; that is, the Laplacian is a symmetric, {\em circulant} matrix defined as

@@

Circulant matrices are diagonalized by the Discrete Fourier Transform \cite{godsil2013algebraic}, and their nonzero eigenvalues come in pairs with multiplicity two, with corresopnding eigenvectors $v_1[n] = 1$, $v_{2m}[n] = \cos(2 \pi mn / N)$, $v_{2m+1}[n] = \sin(2 \pi m n / N)$, $m \geq 1$.  In the case of Equation~\ref{eq:modellaplacian}, it can be shown using the DFT that the eigenvalues are $\lambda_1 = 0$ and

@@

The smallest two nonzero eigenvalues occur when $m = k$, corresponding to the eigenvectors \footnote{This generalizes the circle graph used in \cite{averbuch2015ringit}, in which $T = N, k = 1$} 
\begin{equation}
v_{2k}[n] = \cos(2 \pi n / T), v_{2k+1}[n] = \sin(2 \pi n / T)
\end{equation}
Therefore, the smallest two numerical nonzero eigenvalues each correspond to sinusoids with period $T$ which are mutually orthogonal.  When plotted against each other, they form a circle with an arbitrary phase offset.  Therefore, we compute the circular phase numerically $\phi = \tan^{-1}(\hat{v_1}[n] / \hat{v_2}[n])$, where $\hat{v_i}$ is the eigenvector corresponding to the $i^\text{th}$ smallest nonzero numerical eigenvalue.  In practice, the graphs we get from sliding window may deviate from the ideal Laplacian model in Equation~\ref{eq:modellaplacian}, though they do so gracefully.  Figure~\ref{fig:CirculantExample} shows an example of $\hat{v_1}[n]$ and $\hat{v_2}[n]$ for the unweighted case for an imperfect graph corresponding to the sliding window embedding of a real video.  To further improve robustness, we default to the weighted Laplacian, since it has nice theoretical roperties for Euclidean point placement matching the adjacency matrix \cite{hall1970r}.  In this case, harmonics of the actual frequency of interest occasionally have smaller eigenvalues when the corresponding harmonics are strongly present in the video.  To mitigate this, we use the pair of eigenvectors with the smallest number of zero crossings within the eigenvectors corresponding to the 10 smallest nonzero eigenvalues.

>>>>

Since sliding window embeddings of periodic videos lie on a topological loop, this means that, at an appropriate scale, a graph built on the sliding window point cloud will be approximately circular.  This permits us to compute the {\em phase} $\phi \in [0, 2\pi]$ of each sliding window from this graph, where we use tools from spectral graph theory~\cite{chung1997spectral}, in particular the {\em graph Laplacian}, for phase estimation. We first construct an adjacency matrix $\mb{A}$ in which the entry at $\mb{A}_{ij}$ represents the similarity between windows at times $i$ and $j$.  Given a scale $\sigma$, we define the unweighted adjacency as $\mb{A}_{ij} = 1$ if $||SW_{D}[\mb{Z}(i)] - SW_{D}[\mb{Z}(j)]||_2 \leq \sigma$ and 0 otherwise. We also define a weighted version in which $\mb{A}_{ij} = exp(-||SW_{D}[\mb{Z}(i)] - SW_{D}[\mb{Z}(j)]||_2/2\sigma^2)$. In both cases, we define the graph Laplacian as:
\begin{equation}
	\mb{L} = \mb{D}-\mb{A}
\end{equation}
where $\mb{D}_{ii} = \sum_{j = 1}^N \mb{A}_{ij}, \mb{D}_{i \neq j} = 0$ is the degree matrix representing the sum of all outgoing weights on the diagonal.

We now examine the unweighted Laplacian of an ideal model to motivate our approach.  Suppose we have extracted a sliding window embedding $SW_{D}$ for a video $\mb{Z}$ which repeats itself exactly every $T$ frames $k$ times total, for a total of $N = kT$ windows.  Let us further suppose it is possible to choose an $\sigma$ so that the unweighted adjacency matrix contains an entry of 1 for edges corresponding to windows that are adjacent in time, as well as for windows directly before and after its repetitions, but no others; that is, the Laplacian is a symmetric, {\em circulant} matrix defined as % MSB: "its repetitions": not sure what "its" refers to

@@

Circulant matrices are diagonalized by the Discrete Fourier Transform \cite{godsil2013algebraic}, and their nonzero eigenvalues come in pairs with multiplicity two, with corresponding eigenvectors $v_1[n] = 1$, $v_{2m}[n] = \cos(2 \pi mn / N)$, $v_{2m+1}[n] = \sin(2 \pi m n / N)$, $m \geq 1$.  In the case of Equation~\ref{eq:modellaplacian}, it can be shown using the DFT that the eigenvalues are $\lambda_1 = 0$ and

@@

The smallest two nonzero eigenvalues occur when $m = k$, corresponding to the eigenvectors \footnote{This generalizes the circle graph used in \cite{averbuch2015ringit}, in which $T = N, k = 1$} 
\begin{equation}
v_{2k}[n] = \cos(2 \pi n / T), v_{2k+1}[n] = \sin(2 \pi n / T)
\end{equation}
Therefore, the smallest two numerical nonzero eigenvalues each correspond to sinusoids with period $T$ which are mutually orthogonal.  When plotted against each other, they form a circle with an arbitrary phase offset.  Therefore, we compute the circular phase numerically $\phi = \tan^{-1}(\hat{v_1}[n] / \hat{v_2}[n])$, where $\hat{v_i}$ is the eigenvector corresponding to the $i^\text{th}$ smallest nonzero numerical eigenvalue.

In practice, graphs of the sliding window embedding may deviate from the ideal Laplacian model in Equation~\ref{eq:modellaplacian}, though they do so gracefully.  Figure~\ref{fig:CirculantExample} shows an example of $\hat{v_1}[n]$ and $\hat{v_2}[n]$ for the unweighted case for an imperfect graph corresponding to the sliding window embedding of a real video.  To further improve robustness, we default to the weighted Laplacian, due to its theoretical properties for Euclidean point placement matching the adjacency matrix \cite{hall1970r}.  In this case, harmonics of the actual frequency of interest occasionally have smaller eigenvalues when the corresponding harmonics are strongly present in the video.  To mitigate this, we use the pair of eigenvectors with the smallest number of zero crossings within the eigenvectors corresponding to the 10 smallest nonzero eigenvalues.
% MSB: "real video": what video?
% MSB: what is "point placement"?
